---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rsodb

<!-- badges: start -->
<!-- badges: end -->

rsodb provides an *R* interface to the Spatial Omics database
[SODB][]. The interface is rudimentary, allowing simple discovery and
retrieval of datasets similar to the functionality of the Python
client [pysodb][].

[SODB]: https://gene.ai.tencent.com/SpatialOmics
[pysodb]: https://pysodb.readthedocs.io/en/latest/#

## Installation

Install the development version of rsodb with:

``` r
if (!nzchar(system.file(package = "BiocManager")))
    install.packages("BiocManager", repos = "https://cran.r-project.org")
if (!nzchar(system.file(package = "remotes")))
    BiocManager::install("remotes")
BiocManager::install("mtmorgan/rsodb")
```

## Example

Load the rsodb and dplyr packages

```{r example, message = FALSE}
library(rsodb)
library(dplyr)
```

Discover all available datasets.

```{r discover}
ds <- datasets()
ds
ds |> glimpse()
```

Filter datasets to those of interest using 'standard' dplyr verbs, and
use `local_dataset()` to retrieve experiment results to a local file
cache.

```{r download}
local_dataset <-
    ds |>
    dplyr::filter(dataset == "liu2020high") |>
    local_dataset(dry.run = FALSE)
local_dataset
```

Experiments are **not** downloaded by default; use `dry.run = TRUE` to
download the experiments. Experiments are downloaded only once and
stored in a local cache, use `force = TRUE` to force experiments to be
downloaded again.

Once downloaded, the cached location of individual files are available
with `experiments()`

```{r}
experiments(local_dataset)
```

## Next steps

Files from SODB are in 'AnnData' format, with extension `.h5ad`. These
are easily read into *R* using the [zellkonverter][] package. Make sure it
is installed and attached to the current *R* session (n.b., installing
zellkonverter can be a 'heavy' operation if one is not using
[Bioconductor][], because it depends on many core packages used in
single-cell and spatial omics; the payoff is that one is now ready to
perform advanced analysis single-cell and spatial analysis).

[zellkonverter]: https://bioconductor.org/packages/zellkonverter
[Bioconductor]: https://bioconductor.org

```{r zellkonverter-install}
if (!nzchar(system.file(package = "zellkonverter")))
    BiocManager::install("zellkonverter")
library(zellkonverter)
```

For example, load the first experiment from the 'liu2020high' dataset.

```{r zellkonverter-load, message = FALSE}
local_path <-
    experiments(local_dataset) |>
    ## select the first experiment
    slice(1) |>
    pull("path")
zellkonverter::readH5AD(local_path, use_hdf5 = TRUE)
```

## sessionInfo

The following summarizes software used to render this document.

```{r session-info}
sessioninfo::session_info()
```
